\section{Introduction}

\subsection{Field and Context}

Recent advances in artificial intelligence have led to fully automated workflows that often exceed human performance. State of the art neural networks can classify images into thousands of categories more accurate and magnitudes faster than we can \cite{He2015a}. They translate text between hundreds of languages \cite{Wu2016}, navigate cars autonomously through cities \cite{Bojarski2017} and detect malware in computer systems \cite{Saxe2015}. In most of these cases they have been trained on tens of thousands or even millions of data samples to reach their accuracy. 

Neural networks have also found great success in the medical field, where the requirements are quite different and the data sets are often much smaller. Although the same techniques can be applied, they need to be adjusted to match a particular problem. In many applications neural networks have been setting the current state of the art when it comes to different segmentation or classification tasks.

\subsection{Research Problem}

The age assessment is a challenging procedure to determine the chronological age of a person lacking legal documents. It exists to afford children provisions they are entitled to by law and to prevent adults from taking unwarranted advantage of these benefits. There is no method that offers an exact identification of the age, but several assessments are currently used that can be separated in two groups \cite{EuropeanAsylumSupportOffice2013}.

Personal interviews or examinations are held to gain an observation of physical or psychological features. Besides the fact that these practices include a high amount of manual work comes the problem that they are also subjective to the person conducting the test. The other type of assessments use X-rays to observe physical traits like the collar bone, teeth or carpal growth plates. Besides that these images are also analyzed manually comes the problem that X-rays are considered an invasive imaging method because they require the exposure of ionizing radiation. In some european countries like Germany, these recordings can only be induced in a criminal matter and therefore not be used for refugee applications.

As such, there is a high demand for a fully automated, unprejudiced and non-invasive method for accurately determining the age of a person.

Based on this idea, I was introduced to a data set of MRIs showing the right knee of several candidates. In contrast to CT or X-ray, MRI is considered non-invasive. Similar to the carpal analysis, the three bones around the knee show growth plates which have been used as indicators for the age of a person. Due to the underlying technology used in MRI machines, they are more sensitive to non-bone tissue in comparison to X-ray. Despite this being an advantage is many other medical applications, it adds complex information that is not relevant to the analysis of the bone and growth plates.

Neural networks are known to be feature selectors, meaning that they will learn to extract information out of data samples that are relevant to the task \cite{Setiono1997}. However, this assumes that the data set is large enough to find correlating features, which was not the case based on a series of tests I ran to make age related estimates.

A logical step towards the automated age assessment was therefore removing non-bone tissue from the MR images. The focus of this thesis revolves around creating a fully automated workflow that segments bone in 3D MRIs of human knees. Convolutional neural networks will be the base tool for this study because they have been setting state of the art results in a broad majority of image segmentation tasks.

\subsection{Previous Research}

-----------

The accurate age assessment based on biometric features is a key element in many medical and political

Introduce Field and Context
Summarize previous Research
Define a Research Problem
Introduce the present Research

Statement of the problem: briefly describe problem
Significance of the Study
Objectives of the Study
Time and Place
Scope and Limitation

github link

What is the subject of the paper?
What is the area of interest and what have other researchers found?
How does the current research relate to previous research?
What is the research objective and what hypothesis or research question is being tested?

establish/define any necessary terms/concepts/events/methods etc.
establish your point of view i.e. your main points (what you are going to argue).
orient your reader to what follows i.e. your main points, how you establish your response, the ideas and arguments to be developed.
have a clear and comprehensive statement of your argument in relation to the set topic.

\newpage
